Title

### 2023-04-18 main.tex
Otros cambios

In this section \cpnote{ser coherente con las subsecciones que se tienen nota que no 
hay mencion a los quantum channels. Menos
cuadrado, quizá pasar por chatgpt. hacerlo al final cuanod se tenga el capitulo
por lo menos leido y aprobado en principio de mi parte.} we introduce the
concept of quantum channels, focusing on a specific type called Pauli channels.
Finally, we consider Pauli dynamical maps, which
are curves of Pauli channels parametrized by a variable.

%Quitarrrr:
Then, we propose a quantum circuit that simulates N-qubit Pauli channels
and show the results of implementing the circuit for $N=1$ on
a quantum computer. 
% }}}
\subsection*{Quantum Channels} \label{subsec: Quantum Channels} % {{{

@@

In quantum mechanics, a closed system is represented by a state vector in a 
Hilbert space $\mathcal{H}$. 
The state's evolution is given by Schrodinger's
equation and is always unitary.
However, in real-world situations, 
quantum systems are usually open and interact with their environment in various complicated ways. 
For example, the system's state may become entangled with the environment,
which can lead to loss of information of the system's state over time.

Instead of vectors in the Hilbert space $\mathcal{H}$,
open systems are described by matrices $\rho$ acting on $\mathcal{H}$
called density matrices and they include information about 
the system's interaction with the environment.
To be physically valid, a density matrix $\rho$
has to satisfy the two following conditions:
1. $\tr(\rho) =1$ and 2. $\rho$ is positive semi-definite $(\rho \geq 0)$.

Given that, we can finally define quantum channels, which
describe the time evolution of open quantum systems.
Quantum channels are the most general linear operations
that a quantum system (which is represented 
by a density matrix $\rho$) can undergo independently of its
past~\cite{zimansbook,cirac}.
These channels are constructed based on three fundamental properties: 
linearity, trace preservation, and complete positivity.

Linearity ensures that a quantum channel $\mathcal{E}$ maps any ensemble
of density matrices into the corresponding ensemble of their evolution.
The trace preserving property is given by $\tr (\mathcal{E}[\rho]) = \tr (\rho) =
1$ and guarantees that the quantum channel does not change the condition that $\tr(\rho) = 1$.
Finally, the channel should also preserve
the condition $\rho \geq 0$, and a map that does this is called a positive map.
However, positivity of $\mathcal{E}$ is not enough, 
and we actually require the more restrictive condition of complete positivity.
Complete positivity means that
$\mathcal{E} \otimes \mathbb{I}_n$ is positive for
any positive integer $n$ 
(where $\mathbb{I}_n$ is the $n\times n$ identity matrix).
This is required to ensure that even if $\rho$ is entangled with another system,
applying $\mathcal{E}$ to $\rho$ while doing nothing to the
 other system still results in  a positive semidefinite state.

Given a possible quantum channel $\mathcal{E}$, 
the condition of trace preservation is straightforward to examine but complete positivity is not as simple.
To test complete positivity of a quantum channel, 
Jamiołkowski and Choi~\cite{choi,jamil} developed a simple algorithm.  
The algorithm exploits the
isomorphism between a channel $\mathcal{E}$ and the state $\mathcal{D} =
(\mathbb{I} \otimes \mathcal{E}) [|\Omega \rangle \langle  \Omega|]$, where
$|\Omega\rangle = 1/\text{dim}(\mathcal{H}) \sum_{i}^{\text{dim}(\mathcal{H})} |i \rangle |i \rangle$ is a maximally entangled state between the original system and
an ancilla.  Remarkably, the map $\mathcal{E}$
is completely positive if and only if $\mathcal{D}$ (also known as the Choi or
dynamical matrix of $\mathcal{E}$) is positive semidefinite.

@@

We have discussed the main features of quantum channels and 
now we turn our attention to a specific 
type of channel for $N$-qubit systems called Pauli channels.
First we define a Pauli channel on a one-qubit system,
whose most general single-qubit density
matrix can be written as
\begin{eqnarray}
\label{ec: Density Matrix}
\rho = \dfrac{1}{2} \sum_{\alpha=0}^{3} r_{\alpha} \sigma_{\alpha},
\end{eqnarray}
with $\sigma_0 = \mathbb{I}$, and $\sigma_{1,2,3}$ the usual Pauli matrices.
The condition $\tr(\rho) = 1$ requires that $r_0 = 1$ while $\rho \geq 0$ 
implies that the remaining $r_{1,2,3}$ form
a vector $\vec{r}= (r_1,r_2,r_3)$ inside a unit sphere known as the Bloch sphere. \cpnote{explicar un poco mejor. Me parece que para el publico vale 
la pena ser un poco mas explicito. pero no estoy muy seguro. }. 
A Pauli channel is defined as an operation that
applies each Pauli matrix $\sigma_{\gamma}$ to
the system with probabilities $k_{\gamma}$ with $\gamma = 0,1,2,3$. 
Mathematically, the Pauli channel is written in the following way:
\begin{eqnarray}
\label{ec: Pauli channel 1 qbit}
\mathcal{E}(\rho) = \sum_{\gamma=0}^3 k_{\gamma} \sigma_{\gamma} \rho \sigma_{\gamma},
\end{eqnarray} 
where of course the probabilities $k_{\gamma}$ of applying $\sigma_{\gamma}$
are non-negative real numbers such that 
$\sum_{\gamma} k_{\gamma} = 1$ (these conditions also ensure that the channel is 
trace preserving and completely positive). 
\cpnote{en todo hay que mejorar un poco la redaccion. Aca me gustaria que 
se interpretara. Creo qe es mas que atenuacion de los componentes, o es solo atenuacion? 
en todo caso ver mejor como va la cosa. }

Apart from this interpretation of Pauli channels,
we can find another one by seeing how the channel acts on the 
density matrix of eq. (\ref{ec: Density Matrix}).
To do it, we substitute eq. (\ref{ec: Density Matrix}) in eq. (\ref{ec: Pauli channel 1 qbit}):
\begin{equation}
\mathcal{E}(\rho) = \dfrac{1}{2}\sum_{\gamma,\alpha=0}^3 k_{\gamma} r_{\alpha} \sigma_{\gamma} \sigma_{\alpha} \sigma_{\gamma}.
\end{equation}
This can be simplified by using the following property of Pauli matrices:
\begin{equation}
\label{ec: propiedad-pauli}
\sigma_{\gamma} \sigma_{\alpha} \sigma_{\gamma} = A_{\alpha,\gamma} \sigma_{\alpha}, \;\;\;\;\; \text{with} \; A_{\alpha,\gamma} = \begin{pmatrix}
1 & 1 & 1 & 1\\
1 & 1 & -1 &-1 \\
1 & -1 & 1 & -1 \\
1 & -1 & -1 & 1
\end{pmatrix},
\end{equation}
which leads to
\begin{equation}
\label{ec: rho-transformada}
\mathcal{E}(\rho) = \dfrac{1}{2} \sum_{\alpha} \left(\sum_{\gamma} A_{\alpha, \gamma} k_{\gamma} \right) r_{\alpha} \sigma_{\alpha}.
\end{equation}
Equation \ref{ec: rho-transformada} gives us another interpretation of Pauli channels.
Pauli channels take each component $r_{\alpha}$
of the density matrix and multiply them
by $\sum_{\gamma} A_{\alpha,\gamma} k_{\gamma}$, 
that is:
\begin{equation}
\label{ec: multipliers}
r_{\alpha}  \xrightarrow[\text{Channel}]{\text{Pauli}}  \tau_{\alpha} r_{\alpha} ,\;\; \tau_{\alpha} := \sum_{\gamma} A_{\alpha,\gamma} k_{\gamma}.
\end{equation}
Furthermore, reverting the definition of $\tau_{\alpha}$ by using that $A^{-1} = \dfrac{1}{4} A$, 
we get that $k_{\gamma} = \dfrac{1}{4} \sum_{\gamma} A_{\gamma,\alpha} \tau_{\gamma}$.
Then, using that $k_{\gamma} \geq 0$ and $\sum_{\gamma} k_{\gamma} = 1$,
we get the following conditions on the multipliers:
\begin{eqnarray}
\label{eq: conditions-tetrahedron}
1+\tau_i -\tau_j - \tau_k \geq 0, \\
1+\tau_1 + \tau_2 + \tau_3 \geq 0.
\end{eqnarray}
These conditions imply that $(\tau_1,\tau_2,\tau_3)$
has to be inside a tetrahedron with
vertices $(1,1,1), (1,-1,-1),(-1,1,-1)$ and $(-1,-1,1)$. 
Therefore, the $\tau_{1,2,3}$ are numbers between 
$-1$ and $1$, which means that the components
$r_{\alpha}$ of the density matrix are always 
attenuated and possibly flipped sign.

Pauli channels can be used to model noise....

Having defined the one qubit case, 
in order to present the $N-$qubit one, we 
need to introduce the so-called \textit{Pauli strings}, defined as
\begin{eqnarray}
\label{ec: Pauli string}
\sigma_{\vec{\alpha}} = \sigma_{\alpha_1} \otimes \sigma_{\alpha_2}\otimes \cdots \otimes \sigma_{\alpha_N},
\end{eqnarray}
where $\vec{\alpha}$ denotes a multi-index $(\alpha_1, \cdots, \alpha_N)$
 and $\alpha_i \in \{0,1,2,3\}$. 
These operators form an orthogonal basis in the space of operators acting on $N$ qubits. 
 Similarly to the single-qubit case, the density matrix $\rho$ 
of a system of $N$ qubits can be written using Pauli strings as:
\begin{eqnarray}
\label{ec:  Density matrix Nqbit}
\rho = \dfrac{1}{2^N} \sum_{\vec{\alpha}} r_{\vec{\alpha}} \sigma_{\vec{\alpha}}.
\end{eqnarray}
Then, just as before we define a Pauli channel as a transformation that applies 
the operators $\sigma_{\vec{\gamma}}$ to $\rho$ with probabilities $k_{\vec{\gamma}}$
and is described mathematically by:
\begin{eqnarray}
\label{ec: pauli channel N-qubt}
\mathcal{E}(\rho) = \sum_{\vec{\gamma}} k_{\vec{\gamma}} \sigma_{\vec{\gamma}} \rho \sigma_{\vec{\gamma}},
\end{eqnarray}
where just as before, complete positivity and trace preservation requires
that $k_{\vec{\gamma}}$ are non-negative real numbers such 
that $\sum_{\vec{\gamma}} k_{\vec{\gamma}}=1$.

\textbf{Interpretación de multiplicadores}

\cpnote{Creo qeu toca motivar, conectar con cosas conocidas y en general no ser tan seco}

@@

A Pauli dynamical map is defined as a continuous parametrized 
curve drawn inside the set of Pauli channels and starting at the identity channel. 
Therefore, a Pauli dynamical map can be written as
\begin{eqnarray}
\label{ec: Pauli dynamical map}
\mathcal{E}_p(\rho) = \sum_{\vec{\gamma}} k_{\vec{\gamma}}(p) \sigma_{\vec{\gamma}} \rho \sigma_{\vec{\gamma}},
\end{eqnarray}
where $p$ is a parameter in an interval $[a,b]$ 
and $\mathcal{E}_p$ is a Pauli channel for every $p$, 
with $\mathcal{E}_a$ being the identity channel.

@@

\cpnote{esto está un poco desordenado y a un ritmo creo qeu inapropiado por lo rapido}

We are interested in constructing a quantum circuit that can implement an
arbitrary Pauli channel on a system of $N$ qubits. 
To do it, we remember that a Pauli channel as in eq.~(\ref{ec: pauli channel N-qubt})
is interpreted as a transformation that applies
each Pauli operator $\sigma_{\vec{\gamma}}$ on $\rho$ with a probability $k_{\vec{\gamma}}$.

Given this interpretation of the channel, 
it is easy to construct a quantum circuit that implements 
it with the help of $2n$ ancilla qubits.
The general idea is to create a state in the ancilla qubits that can then
be used to apply the operators $\sigma_{\vec{\gamma}}$
to the system.
The circuit that does this is presented in Fig~(\ref{fig: circuit-pauli}) and its 
first step in the circuit is to create the state
\begin{eqnarray}
\label{ec: state}
\sum_{\vec{\gamma}} b_{\vec{\gamma}} |\vec{\gamma} \rangle
\end{eqnarray}
in the ancilla qubits, 
where $b_{\vec{\gamma}}$ is in general a complex number such that 
\begin{eqnarray}
\label{ec: b}
|b_{\vec{\gamma}}|^2 = k_{\vec{\gamma}}.
\end{eqnarray}
Moreover $|\vec{\gamma}\rangle$ is defined as a $2n$ qubit state given by
\begin{eqnarray}
|\vec{\gamma} \rangle = |\gamma_1\rangle |\gamma_2 \rangle \cdots |\gamma_N \rangle,
\end{eqnarray}
where $|\gamma_i \rangle$ is the $2$ qubit state $|00\rangle, |01\rangle , |10\rangle$ or $|11\rangle$ when $\gamma_i = 0,1,2,3$ respectively. \\

@@

This means that when measured in the computational base, the state given in Eq~(\ref{ec: state})
collapses to $|\vec{\gamma}\rangle$ with a 
probability $|b_{\vec{\gamma}}|^2 = k_{\gamma}$. 
The circuit in Fig~(\ref{fig: circuit-pauli}) uses this fact
to apply the operations $\sigma_{\vec{\gamma}}$ over the principal qubits
with probabilities $k_{\gamma}$ by using controlled operations. 
The symbol shown in Fig~(\ref{fig: control})
indicates that the controlled gate is
activated if these two qubits are in the state $|\gamma_i \rangle$. 
\begin{figure}[h!]
\centering
\begin{quantikz}
& \ctrl{1} & \qw \rstick[wires=2]{$|\gamma_i \rangle$} \\
& \control{} & \qw
\end{quantikz}
\caption{{\bf Control qubits} \tbnote{Al igual que antes, esta imagen se tiene que quitar para la entrega final.} Control $|\gamma_i \rangle$ for two qubits.}
\label{fig: control}
\end{figure}

To see why the circuit from Fig~(\ref{fig: circuit-pauli})
 works, 
we note that after creating the state 
$\sum_{\vec{\gamma}} b_{\gamma} \; |\vec{\gamma} \rangle$, 
the circuit applies each gate $\sigma_{\vec{\gamma}}$ to the principal qubits 
under the condition that the ancilla qubits are 
on the state $|\vec{\gamma}\rangle$. 
This  means that each gate $\sigma_{\vec{\gamma}}$ is applied 
on the principal qubits with a probability $k_{\vec{\gamma}}$, 
just as we wanted.  \\

\subsection*{Simulation for one-qubit Pauli channels} % {{{
\label{subsec: Simulation for one-qubit Pauli channels}
For the particular case of a Pauli channel on one qubit, 
the circuit that simulates it can be constructed as in Fig~(\ref{fig: citcuit-pauli-1}),
 which is a special case of Fig(\ref{fig: circuit-pauli}).

>>>>

In this section \cpnote{ser coherente con las subsecciones que se tienen nota que no 
hay mencion a los quantum channels. Menos
cuadrado, quizá pasar por chatgpt. hacerlo al final cuanod se tenga el capitulo
por lo menos leido y aprobado en principio de mi parte.} we introduce the
concept of quantum channels, focusing on a specific type called Pauli channels.
Furthermore, we define Pauli dynamical maps, which
are curves of Pauli channels parametrized by a variable.
% }}}
\subsection*{Quantum Channels} \label{subsec: Quantum Channels} % {{{

@@

In quantum mechanics, a closed system's state is represented by a vector in a 
Hilbert space $\mathcal{H}$. 
The state's evolution is unitary and given by Schrodinger's equation~\cite{Rieffel}.
However, in real-world situations, quantum systems are usually open, 
which means that they interact with their environment~\cite{Breuer}. 
For instance, the system's state may become entangled with the environment,
leading to a loss of information about the system's state over time.

To describe open systems, instead of state vectors,
we use matrices $\rho$ that act on $\mathcal{H}$.
These matrices are called density matrices, and they include information about the system's interaction with its environment. 
For a density matrix $\rho$ to be physically valid, 
it must satisfy two conditions: $\tr(\rho) = 1$ and $\rho$ is positive semi-definite,
which is denoted as $\rho \geq 0$~\cite{chuangbook}.

Knowing this, we can now define quantum channels.
Quantum channels are operators $\mathcal{E}$ that
describe the evolution of open quantum systems, such that $\rho \rightarrow \mathcal{E}(\rho)$.
Quantum channels are the most general linear operations
that a quantum system can undergo independently of its
past~\cite{zimansbook,cirac}.
These channels are constructed based on three fundamental properties: 
linearity, trace preservation, and complete positivity.

Linearity ensures that a quantum channel $\mathcal{E}$ maps any ensemble
of density matrices into the corresponding ensemble of their evolution.
The trace preserving property is given by $\tr (\mathcal{E}[\rho]) = \tr (\rho) =
1$ and guarantees that the quantum channel does not change the condition that $\tr(\rho) = 1$.
Finally, the channel should also preserve
the condition $\rho \geq 0$, and a map that does this is called a positive map.
However, positivity of $\mathcal{E}$ is not enough, 
and we actually require the more restrictive condition of complete positivity.
Complete positivity means that
$\mathcal{E} \otimes \mathbb{I}_n$ is positive for
any positive integer $n$ 
(where $\mathbb{I}_n$ is the $n\times n$ identity matrix).
This ensures that even if the principal system is entangled with another system,
applying $\mathcal{E}$ to the principal system while doing nothing to the
 other one still results in  a positive semidefinite state for the principal system~\cite{chuangbook}.

Given a quantum channel $\mathcal{E}$, 
the condition of trace preservation is straightforward to verify but complete positivity is not as simple.
To test complete positivity of a quantum channel, 
Jamiołkowski and Choi~\cite{choi,jamil} developed a simple algorithm that exploits the
isomorphism between a channel $\mathcal{E}$ and the state $\mathcal{D} =
(\mathbb{I} \otimes \mathcal{E}) [|\Omega \rangle \langle  \Omega|]$, where
$|\Omega\rangle = 1/\text{dim}(\mathcal{H}) \sum_{i}^{\text{dim}(\mathcal{H})} |i \rangle |i \rangle$ is a maximally entangled state between the original system and
an ancilla.  Remarkably, the map $\mathcal{E}$
is completely positive if and only if $\mathcal{D}$ (also known as the Choi or
dynamical matrix of $\mathcal{E}$) is positive semidefinite.

@@

We have discussed the main features of quantum channels and 
now we turn our attention to a specific 
type of channel for $N$-qubit systems called Pauli channels.
First we will define these channels for single-qubit systems,
whose most general density matrix can be written as~\cite{chuangbook}:
\begin{equation}
\label{eq: Density Matrix}
\rho = \dfrac{1}{2} \sum_{\alpha=0}^{3} r_{\alpha} \sigma_{\alpha},
\end{equation}
with $\sigma_0 = \mathbb{I}$, and $\sigma_{1,2,3}$ the usual Pauli matrices.
The condition $\tr(\rho) = 1$ requires that $r_0 = 1$ while $\rho \geq 0$ 
implies that the remaining $r_{1,2,3}$ form
a vector $\vec{r}= (r_1,r_2,r_3)$ inside a unit sphere known as the Bloch sphere~\cite{Marinescu}.
That is, every possible density matrix for a one-qubit
system is uniquely associated with a point in a unit sphere. \cpnote{explicar un poco mejor. Me parece que para el publico vale 
la pena ser un poco mas explicito. pero no estoy muy seguro.}

Given a one-qubit system described by $\rho$, 
a Pauli channel is defined as an operation that with probability $k_{\gamma}$
applies the Pauli matrix $\sigma_{\gamma}$ to
the system, for $\gamma = 0,1,2,3$~\cite{geometry}.
Mathematically, the Pauli channel is written in the following way:
\begin{eqnarray}
\label{eq: Pauli channel 1 qbit}
\mathcal{E}(\rho) = \sum_{\gamma=0}^3 k_{\gamma} \sigma_{\gamma} \rho \sigma_{\gamma},
\end{eqnarray} 
where of course the probabilities $k_{\gamma}$ of applying $\sigma_{\gamma}$
have to be non-negative real numbers such that 
$\sum_{\gamma} k_{\gamma} = 1$ (these conditions also ensure that the channel is 
trace preserving and completely positive).

Pauli channels are some of the 
most fundamental noise models in quantum information science~\cite{Terhal}. 
Some notable examples of Pauli channels are the following:
\begin{itemize}
\item \textbf{Bit Flip Channel:} This is a channel that with probability $1-p$ leaves the qubit as it is 
and with probability $p$ it applies the $\sigma_1$ matrix 
(which flips the basis states $|0\rangle$ and $|1\rangle$ of the qubit), and
so it is given by:
\begin{align*}
\mathcal{E}(\rho) = (1-p) \rho + p \sigma_1 \rho \sigma_1.
\end{align*}
Analogous channels exist using $\sigma_3$ (called the bit flip channel,
which has a probability $p$ of adding a relative phase $\pi$ to the state) or 
using $\sigma_2$ (called the phase flip channel, 
which has a probability $p$ of 
flipping the base states and also add a relative phase $\pi$).
\item \textbf{Depolarizing channel:} 
This channel has a probability $1-p$
of doing nothing to the qubit and a probability $p$ of converting it into the maximally mixed state $\dfrac{1}{2} \mathbb{I}$
and it can be written as:
\begin{equation}
\mathcal{E}(\rho) = (1-p)\rho + p \dfrac{1}{2} \mathbb{I} = \left(1 - \dfrac{3p}{4} \right) \sigma + \dfrac{p}{4} \sigma_1 \rho \sigma_1 + \dfrac{p}{4} \sigma_2 \rho \sigma_2+ \dfrac{p}{4} \sigma_3 \rho \sigma_3.
\end{equation} 
\end{itemize}

\cpnote{en todo hay que mejorar un poco la redaccion. Aca me gustaria que 
se interpretara. Creo qe es mas que atenuacion de los componentes, o es solo atenuacion? 
en todo caso ver mejor como va la cosa. }

We can also see how an arbitrary Pauli channel acts on an arbitrary  
density matrix.
To do it, we substitute eq. (\ref{eq: Density Matrix}) 
in eq. (\ref{eq: Pauli channel 1 qbit}):
\begin{equation}
\mathcal{E}(\rho) = \dfrac{1}{2}\sum_{\gamma,\alpha=0}^3 k_{\gamma} r_{\alpha} \sigma_{\gamma} \sigma_{\alpha} \sigma_{\gamma}.
\end{equation}
This can be simplified by using the following property of Pauli matrices:
\begin{equation}
\label{eq: propiedad-pauli}
\sigma_{\gamma} \sigma_{\alpha} \sigma_{\gamma} = A_{\alpha,\gamma} \sigma_{\alpha}, \;\;\;\;\; \text{with} \; A_{\alpha,\gamma} = \begin{pmatrix}
1 & 1 & 1 & 1\\
1 & 1 & -1 &-1 \\
1 & -1 & 1 & -1 \\
1 & -1 & -1 & 1
\end{pmatrix},
\end{equation}
which leads to
\begin{equation}
\label{eq: rho-transformada}
\mathcal{E}(\rho) = \dfrac{1}{2} \sum_{\alpha} \left(\sum_{\gamma} A_{\alpha, \gamma} k_{\gamma} \right) r_{\alpha} \sigma_{\alpha}.
\end{equation}
Eq.~(\ref{eq: rho-transformada}) once again has the form of eq.~(\ref{eq: Density Matrix})
but with components $\left( \sum_{\gamma} A_{\alpha,\gamma} k_{\gamma} \right) r_{\alpha}$.
This gives us another way of understanding Pauli channels
as operations that take each component $r_{\alpha}$
of the density matrix and multiplies them
by $\sum_{\gamma} A_{\alpha,\gamma} k_{\gamma}$, 
that is:
\begin{equation}
\label{eq: multipliers}
r_{\alpha}  \xrightarrow[\text{Channel}]{\text{Pauli}}  \tau_{\alpha} r_{\alpha} ,\;\; \tau_{\alpha} := \sum_{\gamma} A_{\alpha,\gamma} k_{\gamma}.
\end{equation}
Notice that $\tau_0 = 1$, which is a consequence of $\sum_{\gamma}k_{\gamma}=1$
and ensures that after the channel, the resulting density matrix still has trace one. 
Furthermore, reverting the definition of $\tau_{\alpha}$ by using that $A^{-1} = \dfrac{1}{4} A$, 
we get that $k_{\gamma} = \dfrac{1}{4} \sum_{\alpha} A_{\alpha,\gamma} \tau_{\alpha}$.
Then, using that $k_{\gamma} \geq 0$
we get the following conditions on the multipliers $\tau_{\alpha}$:
\begin{eqnarray}
\label{eq: conditions-tetrahedron}
&1+\tau_i -\tau_j - \tau_k \geq 0,  \;\;\; \text{for $i,j,k$ different numbers in $\{1,2,3\}$}, \\
&1+\tau_1 + \tau_2 + \tau_3 \geq 0.
\end{eqnarray}

These conditions imply that $(\tau_1,\tau_2,\tau_3)$
has to be inside a tetrahedron with
vertices $(1,1,1), (1,-1,-1),(-1,1,-1)$ and $(-1,-1,1)$. 
Therefore, the $\tau_{1,2,3}$ are numbers between 
$-1$ and $1$, which means that the components
$r_{\alpha}$ of the density matrix are always 
attenuated and possibly sign flipped.

Having defined the one qubit case,  we can now generalize to $N$ qubits.
In order to do it, we need to introduce the so-called \textit{Pauli strings}, defined as
\begin{eqnarray}
\label{eq: Pauli string}
\sigma_{\vec{\alpha}} = \sigma_{\alpha_1} \otimes \sigma_{\alpha_2}\otimes \cdots \otimes \sigma_{\alpha_N},
\end{eqnarray}
where $\vec{\alpha}$ denotes a multi-index $(\alpha_1, \cdots, \alpha_N)$
 and $\alpha_i \in \{0,1,2,3\}$. 
These operators form an orthogonal basis in the space of operators acting on $N$ qubits. 
 Similarly to the single-qubit case, the density matrix $\rho$ 
of a system of $N$ qubits can be written using Pauli strings as:
\begin{equation}
\label{eq:  Density matrix Nqbit}
\rho = \dfrac{1}{2^N} \sum_{\vec{\alpha}} r_{\vec{\alpha}} \sigma_{\vec{\alpha}}.
\end{equation}
Then, just as before, we define a Pauli channel as a transformation that applies 
the operator $\sigma_{\vec{\gamma}}$ to $\rho$ with probability $k_{\vec{\gamma}}$
and is therefore described mathematically by:
\begin{equation}
\label{eq: pauli channel N-qubit}
\mathcal{E}(\rho) = \sum_{\vec{\gamma}} k_{\vec{\gamma}} \sigma_{\vec{\gamma}} \rho \sigma_{\vec{\gamma}},
\end{equation}
where just as before, $k_{\vec{\gamma}}$ are non-negative real numbers such 
that $\sum_{\vec{\gamma}} k_{\vec{\gamma}}=1$.

As in the one qubit case, Pauli channels for $N$ qubits 
attenuate the components $r_{\vec{\alpha}}$ of the density matrix.
This can be seen by substituting eq. (\ref{eq:  Density matrix Nqbit}) in eq. (\ref{eq: pauli channel N-qubit})
and using the property of eq. (\ref{eq: propiedad-pauli}):
\begin{align*}
\mathcal{E}(\rho) &= \dfrac{1}{2^N} \sum_{\vec{\gamma},\vec{\alpha}} k_{\vec{\gamma}} r_{\vec{\alpha}} \sigma_{\vec{\gamma}} \sigma_{\vec{\alpha}} \sigma_{\vec{\gamma}} \\
& = \dfrac{1}{2^N} \sum_{\vec{\alpha}} \left( \sum_{\vec{\gamma}} (A^{\otimes^N})_{\vec{\alpha},\vec{\gamma}} \; k_{\vec{\gamma}} \right) r_{\vec{\alpha}} \sigma_{\vec{\alpha}},
\end{align*}
which means that applying the Pauli channel multiplies the 
components $r_{\vec{\alpha}}$ by $\tau_{\vec{\alpha}} := \sum_{\vec{\gamma}} (A^{\otimes^N})_{\vec{\alpha},\vec{\gamma}} k_{\vec{\gamma}}$.
\cpnote{Creo qeu toca motivar, conectar con cosas conocidas y en general no ser tan seco}

@@

As seen in the last section, Pauli channels and in 
general quantum channels are discrete maps
that transform a density matrix $\rho$ into $\mathcal{E}(\rho)$.
However, we could also define a continuous set of 
channels $\varepsilon_p$ with $p$ a real parameter.

For the special case of Pauli channels, we define 
a Pauli dynamical map as a continuous parametrized 
curve drawn inside the set of Pauli channels and starting at the identity channel. 
Therefore, a Pauli dynamical map can be written as
\begin{eqnarray}
\label{eq: Pauli dynamical map}
\mathcal{E}_p(\rho) = \sum_{\vec{\gamma}} k_{\vec{\gamma}}(p) \sigma_{\vec{\gamma}} \rho \sigma_{\vec{\gamma}},
\end{eqnarray}
where $p$ is a parameter in an interval $[a,b]$ 
and $\mathcal{E}_p$ is a Pauli channel for every $p$, 
with $\mathcal{E}_a$ being the identity channel.

@@

In this section we propose a quantum circuit that 
simulates N-qubit Pauli channels.
Furthermore, we show the results of implementing the circuit for $N=1$ on
a quantum computer. 
\cpnote{esto está un poco desordenado y a un ritmo creo qeu inapropiado por lo rapido}

To create the circuit, we remember that a Pauli channel as in eq.~(\ref{eq: pauli channel N-qubit})
applies each Pauli operator $\sigma_{\vec{\gamma}}$ on $\rho$ with a probability $k_{\vec{\gamma}}$.
Given this, it is easy to construct a quantum circuit that implements 
a Pauli channel with the help of $2N$ ancilla qubits,
the circuit that does this is presented in fig. (\ref{fig: circuit-pauli}).
The general idea is to produce the probabilities $k_{\gamma}$ by creating the state
\begin{eqnarray}
\label{eq: state}
\sum_{\vec{\gamma}} b_{\vec{\gamma}} |\vec{\gamma} \rangle
\end{eqnarray}
in the ancilla qubits, 
where $b_{\vec{\gamma}}$ are numbers such that $
|b_{\vec{\gamma}}|^2 = k_{\vec{\gamma}}.$

%Moreover $|\vec{\gamma}\rangle$ is defined as a $2n$ qubit state given by
%\begin{eqnarray}
%|\vec{\gamma} \rangle = |\gamma_1\rangle |\gamma_2 \rangle \cdots |\gamma_N \rangle,
%\end{eqnarray}
%where $|\gamma_i \rangle$ is the $2$ qubit state $|00\rangle, |01\rangle , |10\rangle$ or $|11\rangle$ when $\gamma_i = 0,1,2,3$ respectively. \\

@@

This means that when measured in the computational basis, the state given in eq.~(\ref{eq: state})
collapses to $|\vec{\gamma}\rangle$ with a 
probability $|b_{\vec{\gamma}}|^2 = k_{\gamma}$. 
The circuit in fig.~(\ref{fig: circuit-pauli}) uses this fact
to operate $\sigma_{\vec{\gamma}}$ on the principal qubits
with a probability $k_{\gamma}$ by using controlled operations 
that apply $\sigma_{\vec{\gamma}}$ to the principal qubits
only if the ancilla qubits are in the state $|\vec{\gamma}\rangle$.
Therefore, the circuit from fig.~(\ref{fig: circuit-pauli}) 
applies each operation $\sigma_{\vec{\gamma}}$ to the principal qubits 
with a probability $k_{\vec{\gamma}}$, 
just as a Pauli channel is supposed to do.  \\

\subsection*{Simulation for one-qubit Pauli channels} % {{{
\label{subsec: Simulation for one-qubit Pauli channels}
For the particular case of a Pauli channel on one qubit, 
the circuit that simulates it can be constructed as in fig. (\ref{fig: citcuit-pauli-1}),
 which is a special case of fig. (\ref{fig: circuit-pauli}).

\begin{figure}[h!]
\centering
\begin{quantikz}
\lstick{$q_0$} & \qw & \gate{\sigma_1} & \gate{\sigma_2} & \gate{\sigma_3} & \qw \\
\lstick{$aq_0$} & \gate[wires=2][3cm]{\quad\quad \text{Create the two qubits state}\quad\quad} & \octrl{-1} & \ctrl{-1} & \ctrl{-1} & \qw \\
\lstick{$aq_1$} & \gateinput{$b_0 |0 \rangle |0 \rangle + b_1 |0 \rangle |1 \rangle + b_2 |1 \rangle |0 \rangle + b_3 |1 \rangle |1 \rangle $} & \ctrl{-2} & \octrl{-2}  & \ctrl{-2} & \qw
\end{quantikz}
\caption{\tbnote{Al igual que antes, hay que adaptar esta figura y quitarla antes de entregar}. {\bf One qubit Pauli channel circuit.} Circuit for a one-qubit Pauli channel, which is a particular case of Fig~(\ref{fig: circuit-pauli}).}
\label{fig: citcuit-pauli-1} 
\end{figure}
Taking advantage of the quantum computers that IBM has available~\cite{Qiskit},
we sampled $249$ one-qubit Pauli channels and simulated them using the 
ibmq-lima quantum computer. 
For each of these channels, we used quantum process tomography to obtain the
Choi matrix corresponding to the implementation of the circuit~\cite{Chuang:1996}.
Then, we compared this Choi matrix with the theoretical
Choi matrix of the channel we were implementing by calculating the 
process fidelity~\cite{Jozsa94,chuangbook} between them.
Finally, using the representation of Pauli channels in a tetrahedron as
in eq.~(\ref{eq: conditions-tetrahedron}),
we show the fidelities for these channels in fig.~(\ref{fig: fidelity one qubit}).
\begin{figure}[h!]
\centering
\includegraphics[width=0.45\textwidth]{fidelity-points.png}\\
\caption{\tbnote{Al igual que antes, hay que quitar esta figura y ponerla en un archivo a parte}. {\bf Fidelities one-qubit Pauli channel} Fidelities of one-qubit Pauli channels implemented on IBM's ibmq-lima. The channels are plotted in the tetrahedron of Pauli channels, with a color bar indicating the fidelity of each channel's implementation using the circuit for a 1-qubit Pauli channel.
\tbnote{Puedo cambiar la fidelidad por otras distancias, como la diamond norm,
pero estoy buscando cuál conviene más.}}w
\label{fig: fidelity one qubit}
\end{figure}

% }}}
% }}}

\newpage
\section*{OPR Circuits} % {{{
\label{sec: OPR Circuits}
Dynamical maps can be implemented using the same circuit as 
in Fig~(\ref{fig: circuit-pauli}), with the only difference that now the state 
to be created on the ancilla qubits depends on parameter $p$ 
and therefore is given by 
\begin{eqnarray}
\label{eq: parametrized state}
\sum_{\vec{\gamma}} b_{\vec{\gamma}}(p) |\vec{\gamma}\rangle.
\end{eqnarray}
